---
title: "Earthquakes"
output: html_notebook
---

I developed a dashboard in Tableau utilizing the earthquake dataset, and now, in my exploration of the data, I am seeking to generate additional visualizations in RStudio to gain further insights and perspectives from the same dataset. This multi-tool approach allows me to leverage the strengths of both Tableau and RStudio for a comprehensive analysis of earthquake patterns and trends.
```{r}
# Install and load necessary packages
install.packages(c("ggplot2", "maps", "readxl"))
library(ggplot2)
library(maps)
library(readxl)
```
```{r}
# GitHub URL of the Excel file
github_url <- "https://raw.githubusercontent.com/DavidWilliamSmith/RStudio/main/Earthquakes/Mag6PlusEarthquakes_1900-2013.xlsx"

# Download the Excel file to a temporary location
temp_file <- tempfile(fileext = ".xlsx")
download.file(github_url, destfile = temp_file, mode = "wb")

# Read the Excel file into R
quakes <- read_excel(temp_file)

# View the first few rows of the dataset
head(quakes)

# Remove the temporary file
unlink(temp_file)
```

```{r}
# Set the background color and font color
theme_custom <- theme(
  plot.background = element_rect(fill = "black"),
  panel.background = element_rect(fill = "black"),
  legend.text = element_text(color = "white"),
  legend.title = element_text(color = "white"),
  axis.text = element_text(color = "white"),
  axis.title = element_text(color = "white")
)
```
# 1. Global Map to display Magnitude Intensity
```{r}
# Create a ggplot with a global map and color contrast for earthquake intensities
gg_global_map <- ggplot() +
  borders("world", colour = "white", fill = NA) +
  geom_point(data = quakes, aes(x = longitude, y = latitude, color = mag), size = 2) +
  scale_color_gradient(low = "lightblue", high = "blue") +
  labs(title = "Global Earthquake Map with Intensity",
       subtitle = "Using ggplot2 in RStudio",
       x = "Longitude",
       y = "Latitude",
       color = "Magnitude") +
  theme_minimal() +
  theme_custom

# Print the global map with color contrast
print(gg_global_map)
```
# 2. Histogram displaying Magnitude Frequencies
```{r}
ggplot(quakes, aes(x = mag)) +
  geom_histogram(binwidth = 0.3, fill = "blue", color = "white", alpha = 0.8) +
  scale_fill_gradient(low = "lightblue", high = "blue") +
  labs(title = "Histogram of Earthquake Magnitudes",
       x = "Magnitude",
       y = "Frequency") +
  theme_minimal() +
  theme_custom

```
# 3. Time Series displaying Earthquakes occurrences in Years
```{r}
# Create a ggplot with a time series of earthquakes by magnitude
gg_time_series <- ggplot(quakes, aes(x = Date, y = ..count.., color = as.factor(mag), group = as.factor(mag))) +
  geom_area(stat = "count", size = 0.5, alpha = 0.7) +
  labs(title = "Time Series of Earthquakes by Magnitude",
       subtitle = "Using ggplot2 in RStudio",
       x = "Date",
       y = "Total Earthquakes",
       color = "Magnitude") +
  scale_color_discrete(name = "Magnitude") +
  theme_minimal() +
  theme_custom

# Print the time series plot
print(gg_time_series)
```





