---
title: "Correlation between Teams Rank and Wins"
output:
  html_document:
    df_print: paged
---

In 2012 and 2013, there were 10 teams in the MLB playoffs: the six teams that had the most wins in each baseball division, and four “wild card” teams. The playoffs start between the four wild card teams - the two teams that win proceed in the playoffs (8 teams remaining). Then, these teams are paired off and play a series of games. The four teams that win are then paired and play to determine who will play in the World Series. 

We can assign rankings to the teams as follows:
 
Rank 1: the team that won the World Series
Rank 2: the team that lost the World Series
Rank 3: the two teams that lost to the teams in the World Series
Rank 4: the four teams that made it past the wild card round, but lost to the above four teams
Rank 5: the two teams that lost the wild card round

In your R console, create a corresponding rank vector by typing
teamRank = c(1,2,3,3,4,4,4,4,5,5)

```{r}
# create a corresponding rank vector
teamRank <- c(1, 2, 3, 3, 4, 4, 4, 4, 5, 5)
```


In this quick question, we’ll see how well these rankings correlate with the regular season wins of the teams. In 2012, the ranking of the teams and their regular season wins were as follows:

Rank 1: San Francisco Giants (Wins = 94)
Rank 2: Detroit Tigers (Wins = 88)
Rank 3: New York Yankees (Wins = 95), and St. Louis Cardinals (Wins = 88)
Rank 4: Baltimore Orioles (Wins = 93), Oakland A’s (Wins = 94), Washington Nationals (Wins = 98), Cincinnati Reds (Wins = 97)
Rank 5: Texas Rangers (Wins = 93), and Atlanta Braves (Wins = 94)

```{r}
# Create a vector in R called wins2012, that has the wins of each team in 2012, in order of rank (the vector should have 10 numbers).
wins2012 <- c(94, 88, 95, 88, 93, 94, 98, 97, 93, 94)
```


In 2013, the ranking of the teams and their regular season wins were as follows:

Rank 1: Boston Red Sox (Wins = 97)
Rank 2: St. Louis Cardinals (Wins = 97)
Rank 3: Los Angeles Dodgers (Wins = 92), and Detroit Tigers (Wins = 93)
Rank 4: Tampa Bay Rays (Wins = 92), Oakland A’s (Wins = 96), Pittsburgh Pirates (Wins = 94), and Atlanta Braves (Wins = 96)
Rank 5: Cleveland Indians (Wins = 92), and Cincinnati Reds (Wins = 90) 



```{r}
# Create another vector in R called wins2013, that has the wins of each team in 2013, in order of rank (the vector should have 10 numbers).
wins2013 <- c(97, 97, 92, 93, 92, 96, 94, 96, 92, 90)
```

What is the correlation between teamRank and wins2012?

# Exercise 1

```{r}
correlation2012 = cor(teamRank, wins2012)
correlation2012
```
# Answer 

teamRanks are 35% positively related to wins2012

What is the correlation between teamRank and wins2013?
# Exercise 2

```{r}
correlation2013 = cor(teamRank, wins2013)
correlation2013
```
# Answer 

teamrRanks are 66% negatively related with wins2013

The correlation coefficient is a statistical measure that quantifies the strength and direction of the linear relationship between two variables. It ranges from -1 to 1, where -1 indicates a perfect negative correlation, 1 indicates a perfect positive correlation, and 0 indicates no linear correlation

A positive correlation coefficient indicates that higher ranks in the playoffs tend to align with higher regular season win totals, suggesting that teams with more wins during the regular season tend to perform better in the playoffs. Conversely, a negative correlation would imply that regular season success does not necessarily translate into playoff success.