---
title: "Getting Started with R,Part 2"
output:
  html_document:
    df_print: paged
---

Let us continue getting started with R as we start discussing important statistical concepts in Sports Analytics.

# Case-scenario 1

This is the fourth season of outfielder Luis Robert with the Chicago White Socks. If during the first three seasons he hit 11, 13, and 12 home runs, how many does he need on this season for his overall average to be at least 20?

# Solution

Given that x1=11,x2=13,x3=12 we want to find x4 such that the mean (average) number of home-runs is x¯>=20

Notice that in this case n=4.

According to the information above: 20×4=11+13+12+x4 so when x4=61, the home-runs average will be 20.
```{r}
# Home-runs so far
HR_before <- c(11, 13, 12)
# Average Number of Home-runs per season wanted
wanted_HR <- 20
# Number of seasons
n_seasons <- 4
# Needed Home-runs on season 4
x_4 <- n_seasons*wanted_HR - sum(HR_before)
# Minimum number of Home-runs needed by Robert
x_4
```

According to the calculations above, Robert must hit 44 home-runs or better on this season to get an average number of home-runs per season of at least 20.

We could confirm this, by using the function mean() in R
```{r}
# Robert's performance
Robert_HRs <- c(11, 13, 12,44)
# Mean
mean(Robert_HRs)
```
```{r}
# Standard Deviation
sd(Robert_HRs)
```
```{r}
# Maximum number of home-runs during the four seasons period
max(Robert_HRs)
```
```{r}
# Minimum number of home-runs during the four seasons period
min(Robert_HRs)
```
```{r}
# Summary 
summary(Robert_HRs)
```

# Question 1

Now, you must complete the problem below which represents a similar case scenario. You may use the steps that we executed in Case-scenario 1 as a template for your solution.

This is the sixth season of outfielder Juan Soto in the majors. If during the first five seasons he received 79, 108,41,145, and 135 walks, how many does he need on this season for his overall number of walks per season to be at least 100?

```{r}
# walks per season so far
walks_per_season <- c(79,108,41,145,135)
# Average Number of Home-runs per season wanted
wanted_walks <- 100
# Number of seasons
n_seasons <- 6
# Needed Home-runs on season 4
x_6 <- n_seasons*wanted_walks - sum(walks_per_season)
# Minimum number of Home-runs needed by Robert
x_6
```
According to the calculations above, Juan must walk 92 times to average 100 walks per season for six seasons

We could confirm this, by using the function mean() in R

```{r}
# Robert's performance
Juans_walks_per_season <- c(79,108,41,145,135,92)
# Mean
mean(Juans_walks_per_season)
```
```{r}
# Standard Deviation
sd(Juans_walks_per_season)
```

```{r}
# Maximum number of home-runs during the four seasons period
max(Juans_walks_per_season)
```

```{r}
# Minimum number of home-runs during the four seasons period
min(Juans_walks_per_season)
```

```{r}
# Summary 
summary(Juans_walks_per_season)
```
# Case-scenario 2

The average salary of 10 baseball players is 72,000 dollars a week and the average salary of 4 soccer players is 84,000. Find the mean salary of all 14 professional players.

# Solution

We can easily find the joined mean by adding both mean and dividing by the total number of people.

Let n1=10 denote the number of baseball players, and y1=72000 their mean salary. 
Let n2=4 the number of soccer players and y2=84000 their mean salary. 
Then the mean salary of all 16 individuals is: n1x1+n2x2n1+n2

We can compute this in R as follows:
```{r}
n_1 <- 10
n_2 <- 4
y_1 <- 72000
y_2 <- 84000
# Mean salary overall
salary_ave <-  (n_1*y_1 + n_2*y_2)/(n_1+n_2)
salary_ave
```

# Question 2

The average salary of 7 basketball players is 102,000 dollars a week and the average salary of 9 NFL players is 91,000. Find the mean salary of all 16 professional players.

# Solution

We can easily find the joined mean by adding both mean and dividing by the total number of people.

Let n1=7 denote the number of baseball players, and y1=102000 their mean salary. 
Let n2=9 the number of soccer players and y2=91000 their mean salary. 
Then the mean salary of all 16 individuals is: n1x1+n2x2n1+n2

We can compute this in R as follows:
```{r}
n_1 <- 7
n_2 <- 9
y_1 <- 102000
y_2 <- 91000
# Mean salary overall
salary_ave <-  (n_1*y_1 + n_2*y_2)/(n_1+n_2)
salary_ave
```

# Case-scenario 3

The frequency distribution below lists the number of active players in the Barclays Premier League and the time left in their contract.

Years	Number of players
6	28
5	72
4	201
3	109
2	56
1	34

Find the mean,the median and the standard deviation.

What percentage of the data lies within one standard deviation of the mean?

What percentage of the data lies within two standard deviations of the mean?

What percent of the data lies within three standard deviations of the mean?

Draw a histogram to illustrate the data.


# Solution

The allcontracts.csv file contains all the players’ contracts length. We can read this file in R using the read.csv() function.
```{r}
contract_length <- read.table("allcontracts.csv", header = TRUE, sep = ",")
contract_years <- contract_length$years
```

Make comments about the code we just ran above.

1.To find the mean and the standard deviation
```{r}
# Mean 
contracts_mean  <- mean(contract_years)
contracts_mean
```
```{r}
# Median
contracts_median <- median(contract_years)
contracts_median
```
```{r}
# Find number of observations
contracts_n <- length(contract_years)
# Find standard deviation
contracts_sd <- sd(contract_years)

contracts_n
contracts_sd
```

2.What percentage of the data lies within one standard deviation of the mean?
```{r}
contracts_w1sd <- sum((contract_years - contracts_mean)/contracts_sd < 1)/ contracts_n
# Percentage of observation within one standard deviation of the mean
contracts_w1sd
```
```{r}
## Difference from empirical 
difference_w1sd <- contracts_w1sd - 0.68
difference_w1sd
```

3. What percentage of the data lies within two standard deviations of the mean?
```{r}
## Within 2 sd
contracts_w2sd <- sum((contract_years - contracts_mean)/ contracts_sd < 2)/contracts_n
contracts_w2sd
```
```{r}
## Difference from empirical 
difference_w2sd <- contracts_w2sd - 0.95
difference_w2sd
```

4. What percent of the data lies within three standard deviations of the mean?
```{r}
## Within 3 sd 
contracts_w3sd <- sum((contract_years - contracts_mean)/ contracts_sd < 3)/contracts_n
contracts_w3sd
```
```{r}
## Difference from empirical 
difference_w3sd <- contracts_w3sd - 0.9973
difference_w3sd 
```

5. Draw a histogram
```{r}
# Create histogram
hist(contract_years,xlab = "Years Left in Contract",col = "green",border = "red", xlim = c(0,8), ylim = c(0,225),  breaks = 5)
```

# Question 3

Use the skills learned in case scenario number 3 on one the following data sets. You may choose only one dataset. They are both available in Canvas.

doubles_hit.csv and triples_hit.csv

Make comments about the code we just ran above.

1.To find the mean and the standard deviation

```{r}
triples_hit <- read.table("triples_hit.csv", header = TRUE, sep = ",")
triples_hit <- triples_hit$triples_hit
```

```{r}
# Mean 
triples_hit_mean  <- mean(triples_hit)
triples_hit_mean
```

```{r}
# Median
triples_hit_median <- median(triples_hit)
triples_hit_median
```

```{r}
# Number of observations
triples_hit_n <- length(triples_hit)
# Standard deviation
triples_hit_sd <- sd(triples_hit)

triples_hit_n
triples_hit_sd
```

2.What percentage of the data lies within one standard deviation of the mean?
```{r}
triples_hit_w1sd <- sum((triples_hit - triples_hit_mean)/triples_hit_sd < 1)/ triples_hit_n
# Percentage of observation within one standard deviation of the mean
triples_hit_w1sd
```
```{r}
## Difference from empirical 
triples_hit_w1sd <- triples_hit_w1sd - 0.68
triples_hit_w1sd
```

3. What percentage of the data lies within two standard deviations of the mean?
```{r}
## Within 2 sd
triples_hit_w2sd <- sum((triples_hit - triples_hit_mean)/ triples_hit_sd < 2)/triples_hit_n
triples_hit_w2sd
```
```{r}
## Difference from empirical 
triples_hit_w2sd <- triples_hit_w2sd - 0.95
triples_hit_w2sd
```

4. What percent of the data lies within three standard deviations of the mean?
```{r}
## Within 3 sd 
triples_hit_w3sd <- sum((triples_hit - triples_hit_mean)/ triples_hit_sd < 3)/triples_hit_n
triples_hit_w3sd
```
```{r}
## Difference from empirical 
triples_hit_w3sd <- triples_hit_w3sd - 0.9973
triples_hit_w3sd 
```

5. Draw a histogram
```{r}
# Create histogram
hist(triples_hit,xlab = "Triples Hit",col = "black",border = "red", xlim = c(0,10), ylim = c(0,50),  breaks = 5)
```