---
title: "First Steps with R"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

# Basic Calculations

You can use R for basic computations you would perform in a calculator
```{r}
#addition
sum <- 2+3
sum
```
```{r}
# division 
divide <- 2/3
divide
```
```{r}
# Exponentiation
expo <- 2^3
expo
```
```{r}
# square root
sqroot <- sqrt(2)
sqroot
```
```{r}
# logarithms
log_2 <- log(2)
log_2
```
```{r}
# compute the log base 5 of 10 
log_base_5 <- log(10, base = 5)
log_base_5

# compute the log of 10
log_10 <- log(10)
log_10
```

Computing some offensive metrics in baseball
```{r}
# Batting Average = (No. of Hits)/(No. of At Bats)
# what is the BA of a player that bats 29 hits in 112 at bats?
BA = (29)/(112)
BA
```
```{r}
Batting_Average=round(BA, digits =3)
Batting_Average
```
```{r}
# what is the BA of a player that bats 42 hits in 212 at bats?
BA2 = (42)/(212)
player2_name_BA = round(BA2, digits = 3)
player2_name_BA
```
Batting average for new player is low at 0.20%

```{r}
#On Base Percentage
#OBP = (H+BB+HBP)/(At Bats +H+BB+HBP+SF)
# lets compute the OBP for a player with the following
#AB = 515,  H = 172, BB = 84, HBP = 5, SF = 6
OBP = (172+84+5)/(515+172+84+5+6)
OBP
```
```{r}
On_Base_Percentage = round(OBP, digits = 3)
On_Base_Percentage
```
```{r}
#OBP = (H+BB+HBP)/(AB+H+BB+HBP+SF)
# compute the OBP for a player with the following general stats:
# AB = 565, H = 156, BB = 65, HBP = 3, SF = 7
OBP2 = (156+65+3)/(565+156+65+3+7)
On_Base_Percentage2 = round(OBP2, digits = 3)
On_Base_Percentage2
```
New OBP is at 0.281%

Often you will want to test whether something is less than, greater than or equal to something
```{r}
3 == 8 # does 3 equal 8?
```
```{r}
3 != 8 # is 3 different from 8?
```
```{r}
3 <= 8 # is 3 less than or equal to 8?
```

The logical operators are & for logical AND, | for logical OR, and ! for NOT. These are some examples:
```{r}
# Logical Disjunction (or)
FALSE | FALSE # False OR False
```
```{r}
# Logical Conjunction (and)
TRUE & FALSE #True AND False
```
```{r}
# Negation
! FALSE # Not False
```
```{r}
# Combination of statements
2 < 3 | 1 == 5 # 2<3 is True, 1==5 is False, True OR False is True
```

# Assigning Values to Variables

In R, you create a variable and assign it a value using <- as follows
```{r}
Total_Bases <- 6 + 5
Total_Bases*3
```

To see the variables that are currently defined, use ls (as in “list”)
```{r}
ls()
```

To delete a variable, use rm (as in “remove”)
```{r}
rm(Total_Bases)
```

Either <- or = can be used to assign a value to a variable, but I prefer <- because is less likely to be confused with the logical operator ==

# Vectors

The basic type of object in R is a vector, which is an ordered list of values of the same type. You can create a vector using the c() function (as in “concatenate”).
```{r}
pitches_by_innings <- c(12, 15, 10, 20, 10) 
pitches_by_innings
```
```{r}
strikes_by_innings <- c(9, 12, 6, 14, 9)
strikes_by_innings
```
```{r}
# Define two vectors,runs_per_9innings and hits_per_9innings, each with five elements. 

runs_per_9innings <- c(2, 2, 3, 2, 1) 
runs_per_9innings

hits_per_9innings <- c(4, 5, 1, 2, 8) 
hits_per_9innings
```

There are also some functions that will create vectors with regular patterns, like repeated elements.
```{r}
# replicate function
rep(2, 5)
```
```{r}
# consecutive numbers
1:5
```
```{r}
# sequence from 1 to 10 with a step of 2
seq(1, 10, by=2)
```

Many functions and operators like + or - will work on all elements of the vector.
```{r}
# add vectors
pitches_by_innings+strikes_by_innings
```
```{r}
# compare vectors
pitches_by_innings == strikes_by_innings
```
```{r}
# find length of vector
length(pitches_by_innings)
```
```{r}
# find minimum value in vector
min(pitches_by_innings)
```
```{r}
# find average value in vector
mean(pitches_by_innings)
```

You can access parts of a vector by using [. Recall what the value is of the vector pitches_by_innings.
```{r}
pitches_by_innings
```
```{r}
# If you want to get the first element:
pitches_by_innings[1]
```
```{r}
# Get the first element of hits_per_9innings?
hits_per_9innings[1]
```
4 hits per inning

If you want to get the last element of pitches_by_innings without explicitly typing the number of elements of pitches_by_innings, make use of the length function, which calculates the length of a vector:
```{r}
pitches_by_innings[length(pitches_by_innings)]
```
```{r}
# Get the last element of hits_per_9innings?
last_element <- hits_per_9innings[length(hits_per_9innings)]
last_element
```
You can also extract multiple values from a vector. For instance to get the 2nd through 4th values use
```{r}
pitches_by_innings[c(2, 3, 4)]
```

Vectors can also be strings or logical values
```{r}
player_positions <- c("catcher", "pitcher", "infielders", "outfielders")
```

# Data Frames

In statistical applications, data is often stored as a data frame, which is like a spreadsheet, with rows as observations and columns as variables.

To manually create a data frame, use the data.frame() function.
```{r}
data.frame(bonus = c(2, 3, 1),#in millions 
           active_roster = c("yes", "no", "yes"), 
           salary = c(1.5, 2.5, 1))#in millions 
```

Most often you will be using data frames loaded from a file. For example, load the results of a fan’s survey. The function load or read.table can be used for this.

# How to Make a Random Sample

To randomly select a sample use the function sample(). The following code selects 5 numbers between 1 and 10 at random (without duplication)
```{r}
sample(1:10, size=5)
```

The first argument gives the vector of data to select elements from.

The second argument (size=) gives the size of the sample to select.

Taking a simple random sample from a data frame is only slightly more complicated, having two steps:

Use sample() to select a sample of size n from a vector of the row numbers of the data frame.
Use the index operator [ to select those rows from the data frame.
Consider the following example with fake data. First, make up a data frame with two columns. (LETTERS is a character vector of length 26 with capital letters âAâ to âZâ; LETTERS is automatically defined and pre-loaded in R)
```{r}
bar <- data.frame(var1 = LETTERS[1:10], var2 = 1:10)
# Check data frame
bar
```

Suppose you want to select a random sample of size 5. First, define a variable n with the size of the sample, i.e. 5
```{r}
n <- 5
```

Now, select a sample of size 5 from the vector with 1 to 10 (the number of rows in bar). Use the function nrow() to find the number of rows in bar instead of manually entering that number.

Use : to create a vector with all the integers between 1 and the number of rows in bar.
```{r}
samplerows <- sample(1:nrow(bar), size=n) 
# print sample rows
samplerows
```

The variable samplerows contains the rows of bar which make a random sample from all the rows in bar. Extract those rows from bar with
```{r}
# extract rows
barsample <- bar[samplerows, ]
# print sample
print(barsample)
```
The code above creates a new data frame called barsample with a random sample of rows from bar.

In a single line of code:
```{r}
bar[sample(1:nrow(bar), n), ]
```

# Using Tables

The table() command allows us to look at tables. Its simplest usage looks like table(x) where x is a categorical variable.

For example, a survey asks people if they support the home team or not. The data is

Yes, No, No, Yes, Yes

We can enter this into R with the c() command, and summarize with the table() command as follows
```{r}
x <- c("Yes","No","No","Yes","Yes") 
table(x)
```

Numerical measures of center and spread
Suppose, MLB Teams’ CEOs yearly compensations are sampled and the following are found (in millions)

12 .4 5 2 50 8 3 1 4 0.25
```{r}
sals <- c(12, .4, 5, 2, 50, 8, 3, 1, 4, 0.25)
# the average
mean(sals) 
```
```{r}
# the variance
var(sals)
```
```{r}
# the standard deviation
sd(sals)
```
```{r}
# the median
median(sals)
```
```{r}
# Tukey's five number summary, usefull for boxplots
# five numbers: min, lower hinge, median, upper hinge, max
fivenum(sals)
```
```{r}
# summary statistics
summary(sals)
```

# How about the mode?

In R we can write our own functions, and a first example of a function is shown below in order to compute the mode of a vector of observations x
```{r}
# Function to find the mode, i.e. most frequent value
getMode <- function(x) {
     ux <- unique(x)
     ux[which.max(tabulate(match(x, ux)))]
 }
```

As an example, we can use the function defined above to find the most frequent value of the number of pitches_by_innings
```{r}
# Most frequent value in baz
getMode(pitches_by_innings)
```


```{r}
# Find the most frequent value of hits_per_9innings using getMode function
most_frequent_value <- getMode(hits_per_9innings)

# Print the most frequent value
print(most_frequent_value)
```


```{r}
# Survey responses 
survey_responses <- c("Saturday", "Saturday", "Sunday", "Monday", "Saturday", "Tuesday", "Sunday", "Friday", "Friday", "Monday")

# Use table() to summarize the survey responses
summary_table <- table(survey_responses)

# Print the summary table
print(summary_table)
```


```{r}
# Use the getMode function to compute the most frequent answer 
most_frequent_answer <- getMode(survey_responses)

# Print the most frequent answer
print(most_frequent_answer)
```

