---
title: "Predicting the Number of Wins"
output:
  html_document:
    df_print: paged
---

```{r}
# Read in data from csv
baseball = read.csv("baseball.csv")
str(baseball)
```
```{r}
# Subset to only include moneyball years
moneyball = subset(baseball, Year < 2002)
str(moneyball)
```
```{r}
# Compute Run Difference
moneyball$RD = moneyball$RS - moneyball$RA
str(moneyball)
```

```{r}
# Scatterplot to check for linear relationship
plot(moneyball$RD, moneyball$W)
```
```{r}
# Regression model to predict wins
WinsReg = lm(W ~ RD, data=moneyball)
summary(WinsReg)
```

```{r}
str(moneyball)
```

```{r}
# Regression model to predict runs scored
RunsReg = lm(RS ~ OBP + SLG + BA, data=moneyball)
summary(RunsReg)
```

```{r}
# Regression model to predict runs scored again but removing the batting average
RunsReg = lm(RS ~ OBP + SLG, data=moneyball)
summary(RunsReg)
```

```{r}
str(moneyball)
```

```{r}
# Regression model to predict runs allowed
RunsAllowedReg = lm(RA ~ OOBP + OSLG, data=moneyball)
summary(RunsAllowedReg)
```

# predicting the number of wins

If a baseball team scores 763 runs and allows 614 runs, how many games do we expect the team to win?
Using the linear regression model constructed during the lecture, enter the number of games we expect the team to win:
   

```{r}
# create run difference for team runs_scored and runs_allowed
runs_scored <- 763
runs_allowed <- 614
run_difference <- runs_scored - runs_allowed
run_difference
```

   
(Intercept) 80.881375   
RD           0.105766

```{r}
intercept <- 80.881375
RD <- 0.105766
wins <- intercept + RD * (run_difference)
wins
```
# Answer

based on the run difference of 149, a team can be predicted to win 97 games.

# predicting the number of runs

# 

If a baseball team’s OBP is 0.361, SLG is 0.409, and BA is 0.257, how many runs do we expect the team to score?
Using the linear regression model constructed during the lecture (the one that uses OBP, SLG, and BA as independent variables), find the number of runs we expect the team to score:

(Intercept)  -788.46      
OBP          2917.42        
SLG          1637.93      
BA           -368.97      

```{r}
intercept_runs <- -788.46      
OBP <- 2917.42
SLG <- 1637.93 
BA <- -368.97

new_OBP <- 0.361
new_SLG <- 0.409
new_BA <- 0.257

runs_scored <- intercept_runs + OBP * (new_OBP) + SLG * (new_SLG) + BA * (new_BA)
runs_scored

runs_scored2 <- -788.46 + 2917.42 * (0.361) + 1637.93 * (0.409) - 368.97 * (0.257)
runs_scored2
```
# Answer  

the team is expected to make 840 runs based on metrics given

#

If a baseball team’s opponents OBP (OOBP) is 0.267 and opponents SLG (OSLG) is 0.392, how many runs do we expect the team to allow?
Using the linear regression model discussed during the lecture (the one on the last slide of the previous video), find the number of runs we expect the team to allow.

(Intercept)  -837.38      
OOBP         2913.60     
OSLG         1514.29    
```{r}
intercept_allowed <-  -837.38      
OOBP <-  2913.60     
OSLG <-  1514.29

new_OOBP <- 0.267
new_OSLG <- 0.392
```

```{r}
runs_allowed <- intercept_allowed + OOBP * (new_OOBP) + OSLG * (new_OSLG)
runs_allowed
```
# Answer

534 runs allowed is predicted for this team

# choosing among different players

Suppose you are the General Manager of a baseball team, and you are selecting two players for your team. You have a budget of $10,500,000, and you have the choice between the following players:

Player Name   OBP   SLG   Salary
Yandy Diaz    0.403 0.511 $8,000,000
Joey Meneses  0.320 0.366 $723,600
Jose Abreu    0.292 0.358 $19,500,000
Ryan Noda     0.384 0.400 $720,000
Nate Lowe     0.365 0.426 $4,050,000

Given your budget and the player statistics, which two players would you select?

# Answer

I would pick Yandy Diaz And Ryan Noda. 
Yandy Diaz would be pick number 1 because he has the highest OBP and SLG of all the players and his salary is below 10mil. To keep the salary cap below 10mil and keep Yandy Diaz, I can choose between Joey Meneses ($724k) and Ryan Noda (720k). Even though the isn't much a difference in the salary between the 2, but their OBP and SLG are noticably different at Joey with OBP = 0.320 and SLG = 0.366 vs Ryan at OBP = 0.384.


